---
title: "Los Angeles Education"
execute: 
  echo: false
---

```{r}
#| warning: false
#| message: false
library(tidyverse)
library(leaflet)
library(sf)
```

```{r}
# Load saved dataset
edu_data <- readRDS("data/edu_data.rds")

la_edu = edu_data |> st_transform(edu_data, crs = 4326)

# Define color palette
pal <- colorNumeric(
  palette = "plasma",
  domain = la_edu$pct_bach_plus,
  na.color = "transparent"
)

# Create labels for hover popups
labels <- sprintf(
  "<strong>%s</strong><br/>%s%% with Bachelor's or higher",
  edu_data$NAME,
  round(la_edu$pct_bach_plus, 1)
) |> lapply(htmltools::HTML)

# Build the leaflet map
leaflet(la_edu) |>
  addProviderTiles("CartoDB.Positron") |>
  addPolygons(
    fillColor = ~pal(pct_bach_plus),
    weight = 0.3,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 1.5,
      color = "#666",
      fillOpacity = 0.8,
      bringToFront = TRUE
    ),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "13px",
      direction = "auto"
    )
  ) |>
  addLegend(
    pal = pal,
    values = ~pct_bach_plus,
    title = "% Bachelor's or Higher",
    position = "bottomright",
    labFormat = labelFormat(suffix = "%", digits = 1)
  )
```

The map illustrates the spatial distribution of educational attainment across census tracts in Los Angeles County, specifically focusing on the percentage of adults with a bachelor's degree or higher. Areas with higher concentrations of college-educated residents—such as Santa Monica, Pasadena, and coastal neighborhoods—appear in warmer colors like yellow and orange, indicating rates above 60%. In contrast, darker purple regions in the northern interior and parts of South and Southeast Los Angeles represent neighborhoods where fewer than 20% of adults hold a bachelor’s degree. The stark geographic disparities highlighted in the map reflect broader socioeconomic divides within the county, suggesting that access to higher education remains unevenly distributed. This visualization provides a powerful tool for identifying communities that may benefit from targeted educational resources or policy interventions.
